---
name: Minnie
surname: Ringland
position: "MESM"
address: "Bren School of Environmental Science & Management"
phone: +1 716 380 7222
email: "mringland@bren.ucsb.edu"
github: mringland
linkedin: minervaringland
date: "`r format(Sys.time(), '%B %Y')`"
output: vitae::moderncv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

# About me

I'm passionate about finding and implementing comprehensive, scalable solutions to the challenges of climate change and pollution. The Master's program at Bren has allowed me to explore policy and economic strategies that promote sustainability while taking into account social and equity implications.

I am interested in
- climate policy
- urban planning
- ESG/CSR
- plastic and waste reduction
- sustainable food systems

# Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "Master of Environmental Science & Management", "Exp. 2021", "Bren School of Environmental Science & Management", "University of California, Santa Barbara",
  "Bachelor of Science, Biology", "2016", "Yale University", "New Haven, Connecticut"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Professional Experience

```{r}
tribble(
  ~Year, ~Type, ~Desc,
  1903, "Physics", "Awarded for her work on radioactivity with Pierre Curie and Henri Becquerel",
  1911, "Chemistry", "Awarded for the discovery of radium and polonium"
) %>% 
  brief_entries(
    glue::glue("Nobel Prize in {Type}"),
    Year, 
    Desc
  )
```

# Publications

```{r}
library(dplyr)
knitr::write_bib(c("vitae", "tibble"), "packages.bib")

bibliography_entries("packages.bib") %>%
  arrange(desc(author$family), issued)
```

